Detecting Facial Expressions in Professional Tennis Matches

### 2017-01-18 Report06-12.Rmd
Analysis Updates

---
title: "Detecting Facial Expressions in Professional Tennis Matches"
author: Stephanie Kobakian, Mitchell O'Hara-Wild. Supervised by Dianne Cook, Stephanie
  Kovalcik
date: "November 19 2016"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    toc_depth: 2
fontsize: 10pt
bibliography: references.bib
---

@@

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE, kfigr.prefix=TRUE)

@@

    }
    if(matchBox){
      p <- p + geom_path(aes(x=x, y=450-y, group = c(type, boxID), colour=boxID), faceData) + guides(colour="none")
    }
    else{
      p <- p + geom_path(aes(x=x, y=450-y, group = type, colour=type), faceData)
    }
  return(p)
  }
}

@@

# Create Venn Diagrams
VennLists<-ALLmetaIMGFaces %>%
  split(.$type) %>% 
  map(~ .$FaceKey)%>% 
  fromList() %>%
  upset(order.by="freq", nsets=100)

@@

##Introduction##

Many tennis professionals believe that tennis is a game heavily affected by the mental states of the players. The opportunity for researching this "inner game" presents itself with the hope of improving the playing and coaching of tennis players by improving their "mental game". 
By statistically analysing the faces and expressions of players during a match there is a hope that insight may be gained into the effects of the mental state on the outcome of a match.
Facial expressions during competition provide the most direct insight into a player's thoughts. The aim of this project is to begin to develop methods to collect accurate information about the facial expressions of elite tennis athletes during matchplay.

In this report, we investigate the performance of several popular facial recognition software's through their Application Programming Interfaces (APIs), and evaluate their performance when applied to the broadcasted videos of elite tennis matches. While it is impossible to know the thoughts and feelings of a player during a match, professionals may be able to infer this information through results produced by a recognition software.

@@

####Project Aim
The major aim of the present study were to determine the feasibility of using currently available facial recognition algorithms for extracting facial information from players during broadcasts of professional matches by comparing the performance of several popular facial recognition APIs. The performance of the evaluated software was compared against manual classification obtained notational tool developed by the authors. In addition to looking at the overall performance, we also evaluated image factors that influences the performance of each service.

####Sample and sampling approach#### 
The goal of the sample was to be representative of the video files that will be used for future facial recognition analysis.

@@

The sample included images that contained the faces of many people, this included players, staff on the court and fans in the crowd. These faces were included in the manual annontations as they were likely to be found by the softwares selected. To be able to contrast the abilities of the softwares and provide information on how to differentiate between players and other people for further research the sample was not filtered at this stage.

@@

####Selecting the software####
The choice of the initial softwares considered for this research were informed by a report that reviewed 'commercial off-the-shelf (COTS) solutions and related patents for face recognition in video surveillance applications.'

The process of software selection to determine which we would compare was based on several criteria. Firstly, we based our choices on the results of the report as it considered processing speed and feature selction techniques, as well as the ability to perform both still-to-video and video-to-video recognition.

From the software's analysed we considered availability for use within the timeframe of the report. This led us to choose Animetrics FaceR. The report outlines that for Animetrics, 'one requirement is that image/face proportion
should be at least 1:8 and that at least 64 pixels between eyes are required'. We realise this could present challenges given our dataset.
It will also allow for an extension from detecting to recognising people in the dataset.

@@

We then chose to consider company's who are expanding their API ranges. This resulted in the choice of Microsoft API, provided by Microsoft Cognitive Services. This detects faces and return a square area where the face was located, and predicts facial features. It also allows the possibility of video stream detection.

@@

We were able to try the online demos to see whether these softwares were viable, we used the following  `r figr("Trial-Image", TRUE, type="Figure")`:

```{r Trial-Image, fig.cap="This is a player", anchor="figure"}
imagesList0<-as.list("2016_HSA_R01_BTomic_AUS_vs_DIstomin_UZB_MS157_clip.0013.png")
overlayGgplot(imagesList0, ALLmetaIMG, matchBox = FALSE)
```

## Methodology ##
An annotation tool was constructed to create a base for comparative analysis, we refer to this as Manual Classifications.
These manual Classifications involved describing the features of the Scenes.

```{r Scene-Attributes-Table, anchor="table", fig.cap="Attributes of a Scene of an image. The most appropriate option from the list was selected for each attribute."}
kable(SceneAttributes, format = "pandoc", booktabs = TRUE)
```

The aim was to collect specific information on each different face within the scene. To determine which of the sometimes many faces in the scene it would be reasonable for softwares to detect a standard was created for reasonable detection. This was based on the attributes provided in `r figr("Solution-Specifications-Table", TRUE, type="Table")`, below:

@@

It was decided that it would be reasonable for a software to detect was decided to be a face size larger than 20 by 20 pixels, as it was specified that the software had minimum distances between the eyes on a face for recognition. (reference table with values)
If it was the face of a player it was recorded if it obviously showed their face. The back of the head was not able to be picked up by any software, so after a demo trial these faces were classified manually  but reclassified as other.
Crowd shots provided difficultly in determining which faces were reasonable to classify. As these faces were not the intended targets of the recognition these faces were contributing to our understanding of the softwares. 
The same face size standard applied to crowd members, but focus was placed on the most prominent faces. For each of these faces, we collected information on the following attributes:

@@

####Manual Annotation
To record the selections of attribute values for each face and scene a Shiny @shiny App was created. We called this Application our ManualClassificationProgram. This helped to provide information for all attributes quickly and consistently.
The program worked by presenting an image, using the imager @imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attibutes in `r figr("Scene-Attributes-Table", TRUE, type="Table")` as shown above.

If there was a face in the image the annotator was able to highlight the square 'Face Box'. This changed the display and presented a set of Attributes with radio buttons, this allowed information to be recorded for the specific face. This recorded the x and y coordinates of the corner points of a box drawn by the mouse, and when the save button was hit it saved all the radio button selections and the 'Face Box' coordinates to a CSV file^[https://github.com/mvparrot/face-recognition/blob/master/ManualClassifiedFaces.csv].

@@

####Software Recognition
The software choices allowed for POST requests to be sent via the internet. To access the APIs through R we enlisted the httr package, using functions from this package a script was written for Google, Animetrics^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/animetrics.R], Microsoft^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/microsoftAPI.R] and Skybiometry^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/autoSkybiometry.R]. These scripts contained loops that would move through the images, individually posting a request for each image to be analysed. These scripts included retrieving the information provided and converting it into a usable format for our analysis.
One interesting anomoly was found when using the Skybiometry software as it limited the amount of requests per minute. We accounted for this by stalling the posts for the amount of waiting time the software notified, and checking until the time lapsed and the script could continue looping.

####Data Processing
The data needed for our analysis was spread across six files. For each software we had the information on the location of the Facial Bounding Boxes, as well as the time taken for the software to find the information. Some of the softwares also provided a more detailed level of information.

@@

A single data set was created to combine all necessary information in the previously mentioned files for our analysis. The information in the data set^[https://github.com/mvparrot/face-recognition/blob/master/ALLmetIMG.csv], was carefully considered. It considers the identify of each face, and all relative face attributes, as well as the image file the face was found in, from this information each face was able to be uniquely identified. Also included was information on the software that found it, and the time it took the software to identify the face. It also has a record of how many faces had been indentified in the image by counting each additional recognised face.
To do so, we gathered the name of the file the face was found in and the software Type the Face Bounding Box was determined by. The automatically determined time values were also included. The minimum and maximum x and y values were drawn from different values in each software's CSV files. This required some processing to align the differing values to be comparable.

To find whether the softwares were recognising the same faces a function was created. As the location and size of the boxes around the faces were recorded, these values were used to see if a particular identified face box matched a manually identified face, or a region found by another software.
This function uses the information of each face and compares the intersecting regions of the polygons created by the x,y coordinates of Manual Faces and other software's faces, to determine if the same face was recognised. We determined the ratio of intersecting area to total area must be greater than 0.1 to be considered the same face. This allowed us to compare the identification areas, as well as contrast the identified faces of each software. This contributed another variable, boxID, to the data set^[https://github.com/mvparrot/face-recognition/blob/master/ALLmetIMG.csv].

####Analysis
The statistical analysis conducted to summarise and assess the validity of the method.
This method allowed all the softwares and the faces they found to be compared.

Using the data set^[ALLmetaIMG] of the combined results, we were able to compare the performance of the softwares.
Firstly, we considered how many individual faces the softwares were able to detect in 
`r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` .

@@

```{r Face-Bounding-Box-Data, results='hide', type="figure", fig.cap="Face Bounding Box Data:", anchor="figure"}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge") +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE)
```

The bar chart in 
`r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` above shows the number of Bounding Boxes produced by each software, comparing the height of the bars indicates that Google's Facial Recognition software recognised almost 1000 more faces than the next best software, Microsoft.

```{r Face-Bounding-Box-Data-Box, anchor="figure"}
ggplot(ALLmetaIMG, aes(x = factor(x = factor(type)), fill=type, y = size)) + geom_boxplot() + scale_y_continuous(trans='log') +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Google","Animetrics","Manual","Microsoft","Skybiometry"))+ylab("Size of Face Bounding Boxes")+guides(fill=FALSE)

@@

The box and whisker plot in 
`r figr("Face-Bounding-Box-Data-Box", TRUE, type="Figure")` shows the differences between the sizes of the 'Face Bounding Boxes' produced by each software. Where a 'Face Bounding Box' refers to the size of the square the software provided as a location in the image of a face that was detected. On average, Google maps the largest boxes around faces, however Animetrics has the largest box recognised in the set. On average, the smallest faces are recognised by Skybiometry.

Skybiometry results accounted for the 255 smallest recognised faces. However this is not necessarily a benefit to this research, as these are not all faces.

```{r I2, results=TRUE, fig.cap="2016_CT6_R02_JSousa_POR_vs_SGiraldo_COL_MS231_clip.0017.png", anchor="figure"}
imagesList2<-as.list("2016_CT6_R02_JSousa_POR_vs_SGiraldo_COL_MS231_clip.0017.png")
overlayGgplot(imagesList2, ALLmetaIMG, matchBox = FALSE)
```

This image shows the smallest Face Bounding Box recognised. However visual inspection shows it only captures the player's eye and nose, not their whole face.

```{r I3, results=TRUE, fig.cap="2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png", anchor="figure"}
imagesList3<-as.list("2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png") 
overlayGgplot(imagesList3, ALLmetaIMG, matchBox = FALSE)
```

As before, the 'face' is not as we would hope it would be classified. Just right of the centre, the smaller box actaully captures a fist, not a face.

```{r I4,results=TRUE, fig.cap="2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png", anchor="figure"}
imagesList4<-as.list("2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png")
overlayGgplot(imagesList4, ALLmetaIMG, matchBox = FALSE)
```

I4 shows the sensitivity of the software. We may consider this too sensitive as it recognised the eye and nose of the ballboy on the court.

##### Image 5.2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png
```{r I5, results=TRUE, fig.cap="2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png", anchor="figure"}
imagesList5<-("2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png") 
overlayGgplot(imagesList5, ALLmetaIMG, matchBox = FALSE)
```

As seen in this image of the crowd, the software is performing reasonably well. It does recognise the faces, yet only one not recognised by the other softwares, this may be due to the sunglasses.

This image demonstrates that the it is not necessarily the smaller faces that Skybiometry recognises, but it puts smaller bounds on the facial features than other softwares.

@@

The sample used contains all the manually annotated faces and all the faces recognised by the four softwares.

To consider how many Type I errors occured, where a face was detected incorrectly, we look at the Bounding Boxes that do not match manually annotated faces.

@@

A potential face detected that does not match a face manually annotated occurs for 9.66% of all the Faces detected by the softwares. This is especially high for Google with 289 faces identified.

All the Face Bounding Boxes that Google found which do not match manually annotated faces were correctly identifying faces. This exhibits the occurences of errors.

We then considered the characteristics of the images that the softwares found Potential Face Bounding Boxes^[These boxes represent an area of pixels that are a potentially recognised face.] in.

@@

`r figr("Image-Characteristics-Table", TRUE, type="Figure")`
 displays the feature combinations that produced the most potential face Bounding Boxes recognitions by all four softwares. The most common shot is a crowd shot.
The second row in the table with 830 faces recognised is more interesting than the first result. This useful scene is an image of a Court Player Close-Up in front of a Logo Wall, taken at Player Shoulder Height.

@@

The best scenes for facial recongition have been found, given this information the following `r figr("Face-Characteristics-Table", TRUE, type="table")`
 considers the Characteristics of the individual faces found within those scenes. For this section we chose to consider only the faces that were manually annotated as players, with the intention of not basing results on recognitions of undesirable faces.

@@

`r figr("Face-Characteristics-Table", TRUE, type="table")`
utilises a set of faces that were Manually annotated and also found by Google. 
Nine of the top ten facial characteristic combinations contained faces that were not wearing glasses.
The head angle describing the face angle was 'Other'^[Definition for headangle factor level 'Other' found in..]
for nine of the top ten facial characteristic combinations.

@@

pander(table(GlassesGoogle$glasses))
```
The extreme disparity between the amount of faces with glasses to without them means that the occurence of these attributes across the softwares must be considered proportionally.

```{r Glasses-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, glasses)
```
Graph 2 above shows that Google outperforms the other softwares, with or without glasses. When there are no glasses worn Google finds over 60% of the manually annotated faces.

@@

pander(table(HeadangleGoogle$headangle))
```
The amount of faces found by Google with the headangle "Other", is much larger than the amount of faces with the headangle "Profile" or "Front On". Therefore this should also be considered propotionately.

```{r Headangle-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, headangle)
```
Graph 3 diplays that Google performs much better in comparison to the other softwares when the headangle is Profile. This is outperforming unusually well, however this could also be due to the poor performance of the softwares in this circumstance.

@@

It should be noted that there were results where a single face was recognised twice within the same image. This was a very unusual result and is notable as a point of interest but given it only occured for Animetrics, it is not worth basing decisions on this unusual result.

@@

`r figr("Google-Characteristics-Table", TRUE, type="table")`
shows how many images with potential player's faces Google recognised. This diplays a vast gap between the amount of potential faces found given the different shot angles. Shoulder height is an optimum angle for a Face Bounding Box.

@@

### Discussion
Graph 1, the Bar Chart of the Face Bounding Boxes promotes Google as the best possible software for a facial recognition application in tennis. According to Table 4 Google had 38.23% of the 9.66% potential face boxes not annotated manually.
It was considered that this may have shown Google's API may have been finding more unwanted faces than the other softwares. However, visual inspection showed that these were actually faces. This is considered in Table 6, which showed that some of the faces were found in a crowd setting. Therefore some of these could be considered irrelevant, possibly being crowd faces. These were deemed unlikely to be detected and neglected during manual annotation. 
Interestingly, Animetrics had the least amount of matches to Manually annotated faces. Looking into this further showed that Animetrics results contained potential Face Bounding Boxes that did not contain faces.

The images were all considered manually. The scene information was recorded and the combinations were shown to find how many potential face Bounding Boxes were found with the combination of scene attributes.
This showed that crowd members faces were often recognised, this is both helpful and unhelpful as it shows a strong ability of Google's algorithm to recognise faces, even when these faces are not the goal of the research.

Image 1 shows the images preferable for future research. Where the faces will be recognised and allow both the identity and emotion of the players to be recognised.

Google gave the optimum results in this image as it found the face of the player, but did not locate the face of the staff memeber behind him on the court.

While Google's recognitions mostly matched the Manually annotated set of faces, there were some that did not. These were all actually faces and were missed during manual annotations.

>>>>

---
title: "Detecting Facial Expressions in Professional Tennis Matches"
author: "Stephanie Kobakian, Mitchell O'Hara-Wild, Dianne Cook, Stephanie Kovalcik"
date: '2017'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    fig_crop: yes
    toc: yes
    toc_depth: 2
  html_notebook: null
fontsize: 11pt
documentclass: article
bibliography: references.bib
---

@@

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE, kfigr.prefix=TRUE, cache=TRUE, fig.env = TRUE)

@@

    }
    if(matchBox){
      p <- p + geom_path(aes(x=x, y=450-y, group = c(type, boxID), colour=boxID), faceData) + guides(colour="none")
    }
    else{
      p <- p + geom_path(aes(x=x, y=450-y, group = c(FaceID), colour=type), faceData)
    }
  return(p)
  }
}

@@

# Create UpSetR graph
createUpSet <- function(data) {
VennLists<-data %>%
  split(.$type) %>% 
  map(~ .$FaceKey)%>% 
  fromList() %>%
  upset(order.by="freq", nsets=100)
}

@@

##Introduction

Many tennis professionals believe that tennis is a game heavily affected by the mental states of the players. The opportunity for researching this "inner game" presents itself with the hope of improving the playing and coaching of tennis players by improving their "mental game". 
By statistically analyzing the faces and expressions of players during a match there is a hope that insight may be gained into the effects of the mental state on the outcome of a match.
Facial expressions during competition provide the most direct insight into a player's thoughts. The aim of this project is to begin to develop methods to collect accurate information about the facial expressions of elite tennis athletes during match play.

In this report, we investigate the performance of several popular facial recognition software's through their Application Programming Interfaces (APIs), and evaluate their performance when applied to the broadcasted videos of elite tennis matches. Using the broadcasted videos gives an objective insight to emotions as the player progresses through a match.
While it is impossible to know the thoughts and feelings of a player during a match, professionals may be able to infer this information through results produced by a recognition software. As opposed to the approach of previous studies that have used Player's recollections after a game to determine their emotions.

@@

####Project Aim
The aims of the present study were to determine the feasibility of using currently available facial recognition algorithms for extracting facial information from players during broadcasts of professional matches by comparing the performance of several popular facial recognition APIs. This limited selection was based on accessible APIs that we believed would produce appropriate and useful facial recognition. The performance of the evaluated software was compared against manual classification obtained notation tool developed by the authors. In addition to looking at the overall performance, we also evaluated image factors that influences the performance of each service.

####Sample and sampling approach
The goal of the sample was to be representative of the video files that will be used for future facial recognition analysis.

@@

The sample included images that contained the faces of many people, this included players, staff on the court and fans in the crowd. These faces were included in the manual annotations as they were likely to be found by the software selected. We felt including these additional faces would not only increase the sample by which to judge the software's capabilities but also allow provision of information on how to differentiate between players and other people for further research. Therefore the sample was not filtered at this initial stage.

@@

####Selecting the software
The choice of the initial software considered for this research were informed by a report that reviewed 'commercial off-the-shelf (COTS) solutions and related patents for face recognition in video surveillance applications.'

The process of software selection to determine which we would compare was based on several criteria. Firstly, we based our choices on the results of the report as it considered processing speed and feature selection techniques, as well as the ability to perform both still-to-video and video-to-video recognition.

From the software's analysed we considered availability for use within the time frame of the report. This led us to choose Animetrics FaceR. The report outlines that for Animetrics, 'one requirement is that image/face proportion
should be at least 1:8 and that at least 64 pixels between eyes are required'. We realize this could present challenges given our data set.
It will also allow for an extension from detecting to recognizing people in the data set.

@@

We then chose to consider companies who are expanding their API ranges. This resulted in the choice of Microsoft API, provided by Microsoft Cognitive Services. This detects faces and return a square area where the face was located, and predicts facial features. It also allows the possibility of video stream detection.

@@

We were able to try the online demos to see whether these software were viable, we used the following Image displayed below in `r figr("Trial-Image", TRUE, type="Figure")`:

```{r Trial-Image, caption="This is a player", message=TRUE, warning=TRUE, anchor="figure"}
imagesList0<-as.list("2016_HSA_R01_BTomic_AUS_vs_DIstomin_UZB_MS157_clip.0013.png")
overlayGgplot(imagesList0, ALLmetaIMG, matchBox = FALSE)
```

## Methodology
An annotation tool was constructed to create a base for comparative analysis, we refer to this as Manual Classifications.
These manual Classifications involved describing the features of the Scenes.

```{r Scene-Attributes-Table, anchor="table", caption="Attributes of a Scene of an image. The most appropriate option from the list was selected for each attribute."}
kable(SceneAttributes, format = "pandoc", booktabs = TRUE)
```

The aim was to collect specific information on each different face within the scene. To determine which of the sometimes many faces in the scene it would be reasonable for software to detect a standard was created for reasonable detection. This was based on the attributes provided in `r figr("Solution-Specifications-Table", TRUE, type="Table")`, below:

@@

It was decided that it would be reasonable for a software to detect was decided to be a face size larger than 20 by 20 pixels, as it was specified that the software had minimum distances between the eyes on a face for recognition. (reference table with values)
If it was the face of a player it was recorded if it obviously showed their face. The back of the head was not able to be picked up by any software, so after a demo trial these faces were classified manually  but reclassified as other.
Crowd shots provided difficultly in determining which faces were reasonable to classify. As these faces were not the intended targets of the recognition these faces were contributing to our understanding of the software. 
The same face size standard applied to crowd members, but focus was placed on the most prominent faces. For each of these faces, we collected information on the following attributes:

@@

####Manual Annotation
To record the details of attributes for each face and scene a Shiny [-@shiny] App was created. We called this Application our ManualClassificationProgram^[https://github.com/mvparrot/face-recognition/blob/master/ManualClassificationProgram.R]. This helped to provide information for all attributes quickly and consistently.
The Shiny [@shinyw] package provides a 'framework for building web applications using R'. It allows for the creation of web applications using R rather than 'requiring HTML, CSS, or JavaScript knowledge.'

The program we created was an interactive application that worked by presenting an image, using the imager @imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attributes in `r figr("Scene-Attributes-Table", TRUE, type="Table")` as shown above.

If there was a face in the image the annotator was able to highlight a section of the image to create a square 'Face Box'. This changed the display and presented a set of Attributes with radio buttons, this allowed information to be recorded for the face in the specific 'Face Box'. This recorded the x and y coordinates of the corner points of a box drawn by the mouse, and when the save button was hit it saved all the radio button selections and the 'Face Box' coordinates to a CSV file^[https://github.com/mvparrot/face-recognition/blob/master/ManualClassifiedFaces.csv].

@@

All the annotations for this sample were completed by one author. This was chosen to provide consistency across the sample of faces annotated manually. However the initial choices of what would be reasonably detected were made by several of the authors.

####Software Recognition
The software choices allowed for POST requests to be sent via the internet. To access the APIs through R we enlisted the httr package, using functions from this package a script was written for Google, Animetrics^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/animetrics.R], Microsoft^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/microsoftAPI.R] and Skybiometry^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/autoSkybiometry.R]. These scripts contained loops that would move through the images, individually posting a request for each image to be analysed. These scripts included retrieving the information provided and converting it into a usable format for our analysis.
One interesting anomaly was found when using the Skybiometry software as it limited the amount of requests per minute. We accounted for this by stalling the posts for the amount of waiting time the software notified, and checking until the time lapsed and the script could continue looping.

####Data Processing
The data needed for our analysis was spread across six files. For each software we had the information on the location of the Facial Bounding Boxes, as well as the time taken for the software to find the information. Some of the software also provided a more detailed level of information.

@@

A single data set was created to combine all necessary information in the previously mentioned files for our analysis. The information in the data set^[https://github.com/mvparrot/face-recognition/blob/master/ALLmetIMG.csv], was carefully considered. It considers the identify of each face, and all relative face attributes, as well as the image file the face was found in, from this information each face was able to be uniquely identified. Also included was information on the software that found it, and the time it took the software to identify the face. It also has a record of how many faces had been identified in the image by counting each additional recognized face.
To do so, we gathered the name of the file the face was found in and the software Type the Face Bounding Box was determined by. The automatically determined time values were also included. The minimum and maximum x and y values were drawn from different values in each software's CSV files. This required some processing to align the differing values to be comparable.

To find whether the software were recognizing the same faces a function was created. As the location and size of the boxes around the faces were recorded, these values were used to see if a particular identified face box matched a manually identified face, or a region found by another software.
This function uses the information of each face and compares the intersecting regions of the polygons created by the x,y coordinates of Manual Faces and other software's faces, to determine if the same face was recognized. We determined the ratio of intersecting area to total area must be greater than 0.1 to be considered the same face. This allowed us to compare the identification areas, as well as contrast the identified faces of each software. This contributed another variable, boxID, to the data set^[https://github.com/mvparrot/face-recognition/blob/master/ALLmetIMG.csv].

####Analysis
The statistical analysis conducted to summarize and assess the validity of the method.
This method allowed all the software and the faces they found to be compared.

Using the data set^[ALLmetaIMG] of the combined API and manual results, we were able to compare the performance of the software.
Firstly, we considered how many individual faces the software were able to detect in `r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` .

However, individual faces are not beneficial if they do not correspond to the faces manually annotated. `r figr("Face-Bounding-Box-UpSet", TRUE, type="Figure")` was created by defining groups depending on the softwares 
that recognized each particular face. The UpSetR [-@UpSetR] package helps visualise set intersections. Where in this circumstance Face Bounding Boxes may overlap on the same face, each bar shows the number of individual faces that have Face Bounding Boxes resulting from each of the softwares highlighted below the bar.

Following, we considered the sizes of the Boxes created in `r figr("Face-Bounding-Box-Data-Box", TRUE, type="Figure")`, this was an attempt to determine how the software performed when presented with smaller faces. The boxplot uses the size of the Face Bounding Boxes, and contrasts based on the software the Boxes were found from.

To determine how each software reacted to the same face we looked at an image that contained only

finding faces that were recognized and grouping them by the softwares

Skybiometry results accounted for the 255 smallest recognized faces. However this is not necessarily a benefit to this research, as these are not all faces.

@@

```{r Face-Bounding-Box-Data, results='hide', type="figure", caption="Face Bounding Box Data:", anchor="figure"}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge") +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE)
```

The bar chart in 
`r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` above shows the number of Bounding Boxes produced by each software, comparing the height of the bars indicates that Google's Facial Recognition software recognized almost 1000 more faces than the next best software, Microsoft.

```{r Face-Bounding-Box-UpSet, results='hide', type="figure", caption="Face Bounding Box Data:", anchor="figure"}
createUpSet(ALLmetaIMGFaces)
```

`r figr("Face-Bounding-Box-Upset", TRUE, type="Figure")` Bar Chart shows the Face Bounding Boxes that were recognized by multiple software.
The largest group with 809 faces is Face Bounding Boxes only found by Manual annotations. The following groups were Boxes recognized by both Manual annotations and the Google API.

```{r Face-Bounding-Box-Data-Box, anchor="figure"}
ggplot(ALLmetaIMG, aes(x = factor(x = factor(type)), fill=type, y = size)) + geom_boxplot() + scale_y_continuous() +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Google","Animetrics","Manual","Microsoft","Skybiometry"))+ylab("Size of Face Bounding Boxes")+guides(fill=FALSE)

@@

The box and whisker plot in 
`r figr("Face-Bounding-Box-Data-Box", TRUE, type="Figure")` shows the differences between the sizes of the 'Face Bounding Boxes' produced by each software. Where a 'Face Bounding Box' refers to the size of the square the software provided as a location in the image of a face that was detected. On average, Google maps the largest boxes around faces, and has the largest box recognized in the set. On average, the smallest Face Boxes are created by Skybiometry.

```{r I2, results=TRUE, caption="2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0014.png", anchor="figure"}
imagesList2<-as.list("2016_CT6_R02_FLopez_ESP_vs_GPella_ARG_MS227_clip.0009.png")
overlayGgplot(imagesList2, ALLmetaIMG, matchBox = FALSE)
```

The image above shows the smallest Face Bounding Box recognized. However visual inspection shows it only captures the player's eye and nose, not their whole face.

```{r I3, results=TRUE, caption="2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png", anchor="figure"}
imagesList3<-as.list("2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png") 
overlayGgplot(imagesList3, ALLmetaIMG, matchBox = FALSE)
```

As before, the 'face' is not as we would hope it would be classified. Just right of the center, the smaller box actually captures a fist, not a face.

```{r I4,results=TRUE, caption="2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png", anchor="figure"}
imagesList4<-as.list("2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png")
overlayGgplot(imagesList4, ALLmetaIMG, matchBox = FALSE)
```

I4 shows the sensitivity of the software. We may consider this too sensitive as it recognized the eye and nose of the ball boy on the court.

```{r I5, results=TRUE, caption="2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png", anchor="figure"}
imagesList5<-("2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png") 
overlayGgplot(imagesList5, ALLmetaIMG, matchBox = FALSE)
```

As seen in this image of the crowd, the software is performing reasonably well. It does recognize the faces, yet only one not recognized by the other software, this may be due to the sunglasses.

This image demonstrates that the it is not necessarily the smaller faces that Skybiometry recognizes, but it puts smaller bounds on the facial features than other software.

@@

The sample used contains all the manually annotated faces and all the faces recognized by the four software.

To consider how many Type I errors occurred, where a face was detected incorrectly, we look at the Bounding Boxes that do not match manually annotated faces.

@@

A potential face detected that does not match a face manually annotated occurs for 9.66% of all the Faces detected by the software. This is especially high for Google with 289 faces identified.

All the Face Bounding Boxes that Google found which do not match manually annotated faces were correctly identifying faces. This exhibits the occurrences of errors.

We then considered the characteristics of the images that the software found Potential Face Bounding Boxes^[These boxes represent an area of pixels that are a potentially recognized face.] in.

@@

`r figr("Image-Characteristics-Table", TRUE, type="Figure")`
 displays the feature combinations that produced the most potential face Bounding Boxes recognition by all four software. The most common shot is a crowd shot.
The second row in the table with 830 faces recognized is more interesting than the first result. This useful scene is an image of a Court Player Close-Up in front of a Logo Wall, taken at Player Shoulder Height.

@@

The best scenes for facial recognition have been found, given this information the following `r figr("Face-Characteristics-Table", TRUE, type="table")`
 considers the Characteristics of the individual faces found within those scenes. For this section we chose to consider only the faces that were manually annotated as players, with the intention of not basing results on recognition of undesirable faces.

@@

`r figr("Face-Characteristics-Table", TRUE, type="table")`
utilizes a set of faces that were Manually annotated and also found by Google. 
Nine of the top ten facial characteristic combinations contained faces that were not wearing glasses.
The head angle describing the face angle was 'Other'^[Definition for head-angle factor level 'Other' found in..]
for nine of the top ten facial characteristic combinations.

@@

pander(table(GlassesGoogle$glasses))
```

The extreme disparity between the amount of faces with glasses to without them means that the occurrence of these attributes across the software must be considered proportionally.

```{r Glasses-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, glasses)
```

Graph 2 above shows that Google outperforms the other software, with or without glasses. When there are no glasses worn Google finds over 60% of the manually annotated faces.

@@

pander(table(HeadangleGoogle$headangle))
```

The amount of faces found by Google with the head-angle "Other", is much larger than the amount of faces with the head-angle "Profile" or "Front On". Therefore this should also be considered proportionately.

```{r Headangle-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, headangle)
```

Graph 3 displays that Google performs much better in comparison to the other software when the head-angle is Profile. This is outperforming unusually well, however this could also be due to the poor performance of the software in this circumstance.

@@

It should be noted that there were results where a single face was recognized twice within the same image. This was a very unusual result and is notable as a point of interest but given it only occurred for Animetrics, it is not worth basing decisions on this unusual result.

@@

`r figr("Google-Characteristics-Table", TRUE, type="table")`
shows how many images with potential player's faces Google recognized. This displays a vast gap between the amount of potential faces found given the different shot angles. Shoulder height is an optimum angle for a Face Bounding Box.

@@

### Discussion
Graph 1, the Bar Chart of the Face Bounding Boxes promotes Google as the best possible software for a facial recognition application in tennis. According to Table 4 Google had 38.23% of the 9.66% potential face boxes not annotated manually.
It was considered that this may have shown Google's API may have been finding more unwanted faces than the other software. However, visual inspection showed that these were actually faces. This is considered in Table 6, which showed that some of the faces were found in a crowd setting. Therefore some of these could be considered irrelevant, possibly being crowd faces. These were deemed unlikely to be detected and neglected during manual annotation. 
Interestingly, Animetrics had the least amount of matches to Manually annotated faces. Looking into this further showed that Animetrics results contained potential Face Bounding Boxes that did not contain faces.

The images were all considered manually. The scene information was recorded and the combinations were shown to find how many potential face Bounding Boxes were found with the combination of scene attributes.
This showed that crowd members faces were often recognized, this is both helpful and unhelpful as it shows a strong ability of Google's algorithm to recognize faces, even when these faces are not the goal of the research.

Image 1 shows the images preferable for future research. Where the faces will be recognized and allow both the identity and emotion of the players to be recognized.

Google gave the optimum results in this image as it found the face of the player, but did not locate the face of the staff member behind him on the court.

While Google's recognition's mostly matched the Manually annotated set of faces, there were some that did not. These were all actually faces and were missed during manual annotations.
